\thesistitle

### 2019-08-24 sections/experimentalsetup.tex
ICSE Results
Experimental setup

\chapter{Experimental Setup}\label{ch:experimentalsetup}
In this chapter we explain the setup of our experiments. All our experiments are quantitative experiments, measured over a corpus using CloneRefactor. In this chapter, we first explain the corpus we use. We then explain the way in which we calculate the impact of refactorings on the software maintainability.

\section{The Corpus}\label{chap:corpus}
For our experiments we use a large corpus of open-source projects assembled by Allamanis et al.~\cite{githubCorpus2013}\footnote{The corpus can be downloaded at: \url{http://groups.inf.ed.ac.uk/cup/javaGithub/}}. This corpus contains a set of Java projects from GitHub, selected by number of forks. The projects and files in this corpus were de-duplicated manually. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

@@

\subsection{Filtering Maven projects}
As explained in Section~\ref{chap:challenge}, CloneRefactor requires all the libraries a software project is dependent on in order to perform its clone analysis. As these are not included in the corpus by Allamanis \cite{githubCorpus2013}, we created a script to gather these dependencies. As Java projects manage their dependencies in many different ways, we decided to limit our scope to a single build automation system. We chose Maven, as it is one of the most used build automation systems for Java projects. Maven uses a file, named \texttt{pom.xml}, to describe a projects' dependencies. Our script clones the latest version of each project in the corpus. It then removes each project that has no \texttt{pom.xml} file (which indicates that those do not use the Maven build automation system).

We then continued to filter all testcases and generated files. As many projects store these files in many different places, we decided to further filter the corpus to only include projects with a conventional structure. For Maven, projects are recommended to put their production code in the folder structure \texttt{src/main/java}. We omitted Maven projects that do not adhere to this convention. We instruct CloneRefactor only to use the sources in the \texttt{src/main/java} folder. This decreases the chance of generated files and testcode being considered for clone detection.

@@

\section{Tool Validation}
We have validated the correctness of CloneRefactor through unit tests and empirical validation. First, we created a set of 57 control projects\footnote{Control projects for testing CloneRefactor: \url{https://github.com/SimonBaars/CloneRefactor/tree/master/src/test/resources}} to verify the correctness in many (edge) cases. These projects test each identified relation, location, contents and refactorability category (see Section \ref{chap:contextsetup} and \ref{sec:refactorabilitysetup}), to see whether they are correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check the correctness of the identified clones, their context and their proposed refactoring.

We also test the correctness of the resulting code after refactoring. For this, we use a GitHub project named JFreeChart. JFreeChart has a high test coverage and working tests, which allows us to test the correctness of the program after running CloneRefactor. We run CloneRefactor manually over the JFreeChart project, run its testcases and check the correctness of the proposed refactorings.

@@

The reason for this is that when applying the ``Extract Method'' refactoring technique, we create a new method. An empty method in Java consists of 7 tokens. We also need calls to the extracted method, which is 4 tokens per call if no arguments are required. Consider the smallest possible clone: 2 clone instances of 1 token. Such a refactoring will never improve maintainability. We reduce duplication by 2 tokens, while increasing volume by 7+4+4+1=16 tokens. This holds true for clones up to 9 tokens (reduce duplication by 18 tokens, increase volume by 7+4+4+9=24 tokens). Empirically, we found all clones below 10 tokens in size, are very unlikely to affect maintainability positively. To avoid such clones polluting our data, we do not consider clones smaller than 10 tokens.

\section{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability (see Section~\ref{sec:metrics}). For each of these metrics, we collect their increase/decrease (delta) after each applied refactoring. We aggregate these delta metric values to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability. % Sander: Dit is niet echt assumptie toch
  \item Normalizing each obtained metric delta over all deltas obtained for a metric in our dataset results in equally weighted scores.
\end{itemize}
We derived these assumptions from supporting evidence shown by Heitlager et al~\cite{heitlager2007practical} and Alves et al.~\cite{alves2010deriving}. Using the resulting aggregated maintainability score, we can argue for each refactoring whether it increases or decreases the maintainability of the system.

We normalize each obtained metric delta using the ``Standard score'', which is calculated as follows:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metric. This method works well for normalization of our data, because as we divide by the standard deviation, outliers do not influence the resulting scores much.

We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
\text{Maintainability Score} = N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}

>>>>

\chapter{Experimental Setup}\label{ch:experimentalsetup}
All our experiments are quantitative experiments, measured over a corpus using CloneRefactor. In this chapter, we first explain the corpus we use. We then explain how we calculate the impact of refactorings on the software maintainability.

\section{The Corpus}\label{chap:corpus}
For our experiments we use a large corpus of open-source projects assembled by Allamanis et al.~\cite{githubCorpus2013}\footnote{The corpus can be downloaded at: \url{http://groups.inf.ed.ac.uk/cup/javaGithub/}}. This corpus contains a set of Java projects from GitHub, selected by the number of forks. The projects and files in this corpus were de-duplicated manually. This results in a variety of Java projects that reflect the quality of average open-source Java systems and are useful to perform measurements on.

@@

\subsection{Filtering Maven projects}
As explained in Section~\ref{chap:challenge}, CloneRefactor requires all the libraries a software project is dependent on to perform its clone analysis. As these are not included in the corpus by Allamanis \cite{githubCorpus2013}, we created a script to gather these dependencies. As Java projects manage their dependencies in many different ways, we decided to limit our scope to a single build automation system. We chose Maven, as it is one of the most used build automation systems for Java projects. Maven uses a file, named \texttt{pom.xml}, to describe a projects' dependencies. Our script clones the latest version of each project in the corpus. It then removes each project that has no \texttt{pom.xml} file (which indicates that those do not use the Maven build automation system).

We then continued to filter all test cases and generated files. As many projects store these files in many different places, we decided to further filter the corpus to only include projects with a conventional structure. For Maven, projects are recommended to put their production code in the folder structure \texttt{src/main/java}. We omitted Maven projects that do not adhere to this convention. We instruct CloneRefactor only to use the sources in the \texttt{src/main/java} folder. This decreases the chance of generated files and test code being considered for clone detection.

@@

\section{Tool Validation}
We have validated the correctness of CloneRefactor through unit tests and empirical validation. First, we created a set of 57 control projects\footnote{Control projects for testing CloneRefactor: \url{https://github.com/SimonBaars/CloneRefactor/tree/master/src/test/resources}} to verify the correctness in many (edge) cases. These projects test each identified relation, location, contents and refactorability category (see Section \ref{chap:contextsetup} and \ref{sec:refactorabilitysetup}), to see whether they are correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check the correctness of the identified clones, their context, and their proposed refactoring.

We also test the correctness of the resulting code after refactoring. For this, we use a GitHub project named JFreeChart. JFreeChart has high test coverage and working tests, which allows us to test the correctness of the program after running CloneRefactor. We run CloneRefactor manually over the JFreeChart project, run its test cases and check the correctness of the proposed refactorings.

@@

The reason for this is that when applying the ``Extract Method'' refactoring technique, we create a new method. An empty method in Java consists of 7 tokens. We also need calls to the extracted method, which is 4 tokens per call if no arguments are required. Empirically, we found all clones below 10 tokens in size, are very unlikely to affect maintainability positively. To avoid such clones polluting our data, we do not consider clones smaller than 10 tokens.

\section{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability (see Section~\ref{sec:metrics}). For each of these metrics, we collect their increase/decrease (delta) after each applied refactoring. We aggregate these delta metric values to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item Normalizing each obtained metric delta over all deltas obtained for that metric in our dataset results in equally weighted scores.
\end{itemize}
We derived these assumptions from supporting evidence shown by Heitlager et al~\cite{heitlager2007practical} and Alves et al.~\cite{alves2010deriving}. Using the resulting aggregated maintainability score, we can argue for each refactoring whether it increases or decreases the maintainability of the system.

We normalize each obtained metric delta using the ``Standard score'', which is calculated as follows:
\begin{equation}\label{eq:scoredev}
N_{metric} = \frac {\Delta X-\mu}{\sigma}
\end{equation}
Where $\Delta X$ is a metric delta, $\mu$ is the mean of all deltas for this metric and $\sigma$ is the standard deviation of all deltas for this metric. This method works well for normalization of our data because it is resilient against outliers. We divide by the standard deviation so outliers do not influence the resulting scores much.

We then calculate the maintainability score for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
\text{Maintainability Score} = N_{duplication} + N_{complexity} + N_{volume} + N_{parameters}
\end{equation}
